---
title: 'The C-statistic provides no information on the accuracy of program evaluations:
  Simulation Code'
author: "Andrew Wey"
date: "Spring 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setting up c_effect_est_params

The primary function (sim_c_effect_iter_fun) for simulating one iteration requires c_effect_est_params to create a covariate distribution and survival times representative of deceased-donor kidney-alone model from the January 2018 program-specific reports.  We construct c_effect_est_params from files in gitlab; note that the variance parameters were estimated from patient-level data.


### Importing variance components of the linear predictors

```{r lp_var_components}

lp_var_components <- read.csv("lp_component_var.csv",
                              header = FALSE,
                              stringsAsFactors = FALSE)

```


### Importing vector of program sizes

```{r program_sizes}

program_sizes <- read.csv("ctr_n_vec.csv",
                          header = FALSE,
                          stringsAsFactors = FALSE)

```


### Importing variance for program-level hazard ratios

```{r program_var}

program_var <- read.csv("program_var.csv",
                        header = FALSE,
                        stringsAsFactors = FALSE)

```


### Importing information on baseline hazard function

```{r bl_hazard_info}

bl_hazard_info <- read.csv("bl_hazard.csv",
                           header = TRUE,
                           stringsAsFactors = FALSE)

```


### Creating c_effect_est_params

```{r create_c_effect_est_params}

c_effect_est_params <- list(var_components = unlist(lp_var_components),
                            program_size = unlist(program_sizes),
                            program_var = unlist(program_var),
                            baseline_hazard = bl_hazard_info)

```




## Run Simulations

As a warning, the simulations will likely take several hours to complete.


### Source necessary files

```{r source_sim_files}

library(survival)
library(dplyr)

source("sim_c_effect_iter_fun.R")
source("gen_surv_time.R")

```


### Create directory for simulations

Due to the relatively high computational resources, the simulations are saved to disk.


```{r create_sim_dir}

### This may create issues

if(! dir.exists(file.path(getwd(), "/simulations"))){ dir.create(file.path(getwd(), "/simulations")) }

```


### Run and write simulation results

```{r run_c_effect_sims}

if(! file.exists("simulations/c_effect_lp_8_ctrvar_1.5_confound_third.Rdata")){

  c_effect_lp_0.5_ctrvar_1 <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 0.5, 1) })
  c_effect_lp_1_ctrvar_1   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 1, 1) })
  c_effect_lp_2_ctrvar_1   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 2, 1) })
  c_effect_lp_4_ctrvar_1   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 4, 1) })
  c_effect_lp_8_ctrvar_1   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 8, 1) })

  save(c_effect_lp_0.5_ctrvar_1, file = "simulations/c_effect_lp_0.5_ctrvar_1.Rdata")
  save(c_effect_lp_1_ctrvar_1, file = "simulations/c_effect_lp_1_ctrvar_1.Rdata")
  save(c_effect_lp_2_ctrvar_1, file = "simulations/c_effect_lp_2_ctrvar_1.Rdata")
  save(c_effect_lp_4_ctrvar_1, file = "simulations/c_effect_lp_4_ctrvar_1.Rdata")
  save(c_effect_lp_8_ctrvar_1, file = "simulations/c_effect_lp_8_ctrvar_1.Rdata")

  
  
  c_effect_lp_0.5_ctrvar_1.5_confound_third <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 0.5, 1.5, 1/3) })
  c_effect_lp_1_ctrvar_1.5_confound_third   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 1, 1.5, 1/3) })
  c_effect_lp_2_ctrvar_1.5_confound_third   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 2, 1.5, 1/3) })
  c_effect_lp_4_ctrvar_1.5_confound_third   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 4, 1.5, 1/3) })
  c_effect_lp_8_ctrvar_1.5_confound_third   <- sapply(1:1000, function(x){ sim_c_effect_iter_fun(x, 8, 1.5, 1/3) })

  save(c_effect_lp_0.5_ctrvar_1.5_confound_third, file = "simulations/c_effect_lp_0.5_ctrvar_1.5_confound_third.Rdata")
  save(c_effect_lp_1_ctrvar_1.5_confound_third, file = "simulations/c_effect_lp_1_ctrvar_1.5_confound_third.Rdata")
  save(c_effect_lp_2_ctrvar_1.5_confound_third, file = "simulations/c_effect_lp_2_ctrvar_1.5_confound_third.Rdata")
  save(c_effect_lp_4_ctrvar_1.5_confound_third, file = "simulations/c_effect_lp_4_ctrvar_1.5_confound_third.Rdata")
  save(c_effect_lp_8_ctrvar_1.5_confound_third, file = "simulations/c_effect_lp_8_ctrvar_1.5_confound_third.Rdata")
}

```



## Summarize simulation results

### Import simulation results and average over the iterations

```{r summarize_c_effect_sims}


ctrvar_1_results          <- NULL
ctrvar_1_confound_results <- NULL


pxvar <- c(0.5, 1, 2, 4, 8)


for(i in pxvar){
  
  ### Load and bind the ctrvar 1 component
  load(file = file.path("simulations", 
                        paste0("c_effect_lp_", 
                               i,
                               "_ctrvar_1.Rdata")))
  
  ctrvar_1_results <- rbind(ctrvar_1_results,
                            apply(get(paste0("c_effect_lp_", 
                                             i,
                                             "_ctrvar_1")),
                                  1, 
                                  mean,
                                  na.rm = TRUE))
  
  
  
  ### Load and bind the ctrvar 1 component with unmeasured confounding
  load(file = file.path("simulations", 
                        paste0("c_effect_lp_", 
                               i,
                               "_ctrvar_1.5_confound_third.Rdata")))
  
  ctrvar_1_confound_results <- rbind(ctrvar_1_confound_results,
                                     apply(get(paste0("c_effect_lp_", 
                                                      i,
                                                      "_ctrvar_1.5_confound_third")),
                                           1, 
                                           mean,
                                           na.rm = TRUE))
}

```



### Create objects to more easily summarize results

```{r create_summary_objs}

empirical_c_stats <- cbind(ctrvar_1_results[,1],
                           ctrvar_1_confound_results[,1])


empirical_expected <- cbind(ctrvar_1_results[,2],
                            ctrvar_1_confound_results[,2])


empirical_1yr_surv <- cbind(ctrvar_1_results[,3],
                            ctrvar_1_confound_results[,3])


surv_0.75_qnt <- cbind(ctrvar_1_results[,4],
                       ctrvar_1_confound_results[,4])


surv_0.90_qnt <- cbind(ctrvar_1_results[,5],
                       ctrvar_1_confound_results[,5])


surv_0.95_qnt <- cbind(ctrvar_1_results[,6],
                       ctrvar_1_confound_results[,6])


surv_0.99_qnt <- cbind(ctrvar_1_results[,7],
                       ctrvar_1_confound_results[,7])


shr_exp_1   <- ctrvar_1_results[, 2:4 * 19 - 11]
shr_bias_1  <- ctrvar_1_results[, 2:4 * 19 - 8]
shr_mse_1   <- ctrvar_1_results[, 2:4 * 19 - 6]
shr_shr_1   <- ctrvar_1_results[, 2:4 * 19 - 4]
shr_5tier_1 <- ctrvar_1_results[, 2:4 * 19 - 2]
shr_HRt1_1  <- ctrvar_1_results[, 2:4 * 19 + 1]
shr_HRt2_1  <- ctrvar_1_results[, 2:4 * 19 + 2]
shr_HRt3_1  <- ctrvar_1_results[, 2:4 * 19 + 3]
shr_HRt4_1  <- ctrvar_1_results[, 2:4 * 19 + 4]
shr_HRt5_1  <- ctrvar_1_results[, 2:4 * 19 + 5]
shr_cms_1   <- ctrvar_1_results[, 2:4 * 19 + 6]
shr_mpsc_1  <- ctrvar_1_results[, 2:4 * 19 + 7]



shr_exp_1_confound   <- ctrvar_1_confound_results[, 2:4 * 19 - 11]
shr_bias_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 - 8]
shr_mse_1_confound   <- ctrvar_1_confound_results[, 2:4 * 19 - 6]
shr_shr_1_confound   <- ctrvar_1_confound_results[, 2:4 * 19 - 4]
shr_5tier_1_confound <- ctrvar_1_confound_results[, 2:4 * 19 - 2]
shr_HRt1_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 1]
shr_HRt2_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 2]
shr_HRt3_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 3]
shr_HRt4_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 4]
shr_HRt5_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 5]
shr_cms_1_confound   <- ctrvar_1_confound_results[, 2:4 * 19 + 6]
shr_mpsc_1_confound  <- ctrvar_1_confound_results[, 2:4 * 19 + 7]


```



### Summary statistics (Table 2)


```{r ctrvar_1_characteristics}


desc_table <- cbind(empirical_c_stats[,1],
                    empirical_expected[,1],
                    empirical_1yr_surv[,1],
                    surv_0.75_qnt[,1],
                    surv_0.90_qnt[,1],
                    surv_0.95_qnt[,1],
                    surv_0.99_qnt[,1])


desc_table_nice <- cbind(format(round(desc_table[,1], 2), nsmall = 2),
                         round(desc_table[,2]),
                         paste0(format(round(100 * desc_table[,3], 1), nsmall = 1), "%"),
                         paste0(format(round(100 * desc_table[,4], 1), nsmall = 1), "%"),
                         paste0(format(round(100 * desc_table[,5], 1), nsmall = 1), "%"),
                         paste0(format(round(100 * desc_table[,6], 1), nsmall = 1), "%"),
                         paste0(format(round(100 * desc_table[,7], 1), nsmall = 1), "%"))


dimnames(desc_table_nice) <- list(NULL,
                                  c("C-Statistic",
                                    "Expected Events",
                                    "Observed",
                                    paste0(c(75, 90, 95, 99),
                                           "th percentile")))


print(desc_table_nice)


```






### Comparison without and with unadjusted risks (Table 3)


```{r unadj_risk_comparison}


unadj_risks_table <- cbind(empirical_c_stats,
                           empirical_expected,
                           empirical_1yr_surv)



unadj_risks_table_nice <- cbind(format(round(unadj_risks_table[,1], 2), nsmall = 2),
                                format(round(unadj_risks_table[,2], 2), nsmall = 2),
                                round(unadj_risks_table[,3]),
                                round(unadj_risks_table[,4]),
                                paste0(format(round(100 * unadj_risks_table[,5], 1), nsmall = 1), "%"),
                                paste0(format(round(100 * unadj_risks_table[,6], 1), nsmall = 1), "%"))




dimnames(unadj_risks_table_nice) <- list(NULL,
                                         paste0(rep(c("C-statistic",
                                                     "Expected",
                                                     "1-year survival"), 
                                                    each = 2),
                                                ": ",
                                               rep(c("Without",
                                                     "With"),
                                                   3)))

print(unadj_risks_table_nice)


```




## Mean-squared error (Figure 1)

```{r MSE_figure}


par(
  bty = 'n')

par(mfrow = c(1,2))


plot(shr_mse_1[,1] ~ empirical_c_stats[,1],
     xlab = "C-statistic",
     ylab = "Mean-Squared Error (MSE)",
     ylim = c(0, 0.20),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
mtext("No unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_mse_1[,2] ~ empirical_c_stats[,1],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_mse_1[,3] ~ empirical_c_stats[,1],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)





plot(shr_mse_1_confound[,1] ~ empirical_c_stats[,2],
     xlab = "C-statistic",
     ylab = "Mean-Squared Error (MSE)",
     ylim = c(0, 0.20),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
mtext("Unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_mse_1_confound[,2] ~ empirical_c_stats[,2],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_mse_1_confound[,3] ~ empirical_c_stats[,2],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)
legend("bottomleft",
       c("<3",
         "3-<10",
         expression("" >= "10")),
       lty = c(1, 2, 5),
       pch = c(6, 2, 5),
       lwd = c(2, 2, 2),
       bty = "n",
       title = "Expected Events")


```


## Five-tier (Figure 2)


```{r fivetier_figure}


par(
  bty = 'n')

par(mfrow = c(1,2))


plot(shr_5tier_1[,1] ~ empirical_c_stats[,1],
     xlab = "C-statistic",
     ylab = "Kendall's tau",
     ylim = c(-1, 0),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
mtext("No unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_5tier_1[,2] ~ empirical_c_stats[,1],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_5tier_1[,3] ~ empirical_c_stats[,1],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)





plot(shr_5tier_1_confound[,1] ~ empirical_c_stats[,2],
     xlab = "C-statistic",
     ylab = "Kendall's tau",
     ylim = c(-1, 0),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
mtext("Unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_5tier_1_confound[,2] ~ empirical_c_stats[,2],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_5tier_1_confound[,3] ~ empirical_c_stats[,2],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)
legend("bottomleft",
       c("<3",
         "3-<10",
         expression("" >= "10")),
       lty = c(1, 2, 5),
       pch = c(6, 2, 5),
       lwd = c(2, 2, 2),
       bty = "n",
       title = "Expected Events")


```



## Flag-Status (Figure 3)


```{r flagstatus_figure}


par(
  bty = 'n')

par(mfrow = c(2,2))


plot(shr_cms_1[,1] ~ empirical_c_stats[,1],
     xlab = "C-statistic",
     ylab = "Probability",
     ylim = c(0.5, 1),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
title(main = "CMS Flag Status",
      line = 3)
mtext("No unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_cms_1[,2] ~ empirical_c_stats[,1],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_cms_1[,3] ~ empirical_c_stats[,1],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)
legend("bottomright",
       c("<3",
         "3-<10",
         expression("" >= "10")),
       lty = c(1, 2, 5),
       pch = c(6, 2, 5),
       lwd = c(2, 2, 2),
       bty = "n",
       title = "Expected Events")





plot(shr_cms_1_confound[,1] ~ empirical_c_stats[,2],
     xlab = "C-statistic",
     ylab = "Probability",
     ylim = c(0.5, 1),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
title(main = "CMS Flag Status",
      line = 3)
mtext("Unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_cms_1_confound[,2] ~ empirical_c_stats[,2],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_cms_1_confound[,3] ~ empirical_c_stats[,2],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)





plot(shr_mpsc_1[,1] ~ empirical_c_stats[,1],
     xlab = "C-statistic",
     ylab = "Probability",
     ylim = c(0.5, 1.0),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
title(main = "MPSC Flag Status",
      line = 3)
mtext("No unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_mpsc_1[,2] ~ empirical_c_stats[,1],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_mpsc_1[,3] ~ empirical_c_stats[,1],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)





plot(shr_mpsc_1_confound[,1] ~ empirical_c_stats[,2],
     xlab = "C-statistic",
     ylab = "Probability",
     ylim = c(0.5, 1.0),
     xlim = c(0.5, 1.0),
     type = "b",
     pch = 6,
     lwd = 2,
     main = "")
title(main = "MPSC Flag Status",
      line = 3)
mtext("Unadjusted risk factors",
      side = 3,
      line = 1)
lines(shr_mpsc_1_confound[,2] ~ empirical_c_stats[,2],
      type = "b",
      pch = 2,
      lty = 2,
      lwd = 2)
lines(shr_mpsc_1_confound[,3] ~ empirical_c_stats[,2],
      type = "b",
      pch = 5,
      lty = 5,
      lwd = 2)
legend("topleft",
       c("<3",
         "3-<10",
         expression("" >= "10")),
       lty = c(1, 2, 5),
       pch = c(6, 2, 5),
       lwd = c(2, 2, 2),
       bty = "n",
       title = "Expected Events")


```












